---
title: "Billionaires"
author: "Bright Amenyo (5520368)"
date: "2023-12-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Packages 
library(dplyr) # Glimpse
```
Attributes 
rank: The ranking of the billionaire in terms of wealth.
finalWorth: The final net worth of the billionaire in U.S. dollars.
category: The category or industry in which the billionaire's business operates.
personName: The full name of the billionaire.
age: The age of the billionaire.
country: The country in which the billionaire resides.
city: The city in which the billionaire resides.
source: The source of the billionaire's wealth.
industries: The industries associated with the billionaire's business interests.
countryOfCitizenship: The country of citizenship of the billionaire.
organization: The name of the organization or company associated with the billionaire.
selfMade: Indicates whether the billionaire is self-made (True/False).
status: "D" represents self-made billionaires (Founders/Entrepreneurs) and "U" indicates inherited or unearned wealth.
gender: The gender of the billionaire.
birthDate: The birthdate of the billionaire.
lastName: The last name of the billionaire.
firstName: The first name of the billionaire.
title: The title or honorific of the billionaire.
date: The date of data collection.
state: The state in which the billionaire resides.
residenceStateRegion: The region or state of residence of the billionaire.
birthYear: The birth year of the billionaire.
birthMonth: The birth month of the billionaire.
birthDay: The birth day of the billionaire.
cpi_country: Consumer Price Index (CPI) for the billionaire's country.
cpi_change_country: CPI change for the billionaire's country.
gdp_country: Gross Domestic Product (GDP) for the billionaire's country.
gross_tertiary_education_enrollment: Enrollment in tertiary education in the billionaire's country.
gross_primary_education_enrollment_country: Enrollment in primary education in the billionaire's country.
life_expectancy_country: Life expectancy in the billionaire's country.
tax_revenue_country_country: Tax revenue in the billionaire's country.
total_tax_rate_country: Total tax rate in the billionaire's country.
population_country: Population of the billionaire's country.
latitude_country: Latitude coordinate of the billionaire's country.
longitude_country: Longitude coordinate of the billionaire's country.

#LOADING DATASET
```{r}
Billion_data<-read.csv("Billionaires.csv",header = TRUE)

# Display the structure of the dataset
glimpse(Billion_data) ## str(Billion_data)


```


#Cleaning Dataset
```{r}
# Convert empty strings to NA
Billion_data[Billion_data == ""] <- NA ## Just to make sure all the blanks spaces are NA

# Count missing values in each column
missing_values <- colSums(is.na(Billion_data))
# Create a data frame with column names and their corresponding missing counts
missing_df <- data.frame(column = names(missing_values), missing_count = missing_values)
# Sort the data frame by missing counts in descending order
missing_df <- missing_df[order(-missing_df$missing_count), ]

# Display the result
print(missing_df)

```
# Handling missimg observations
-Organization
```{r}

# Wild guess: assuming that when there's missing value under the organization section then it was unearned or inherited
# Print specific columns
print(Billion_data[c('organization', 'selfMade')])

# Print the distribution of values in the 'title' column
title_counts <- table(Billion_data$title)

# Replace missing values in 'organization' when 'SelfMade' is 'FALSE'
Billion_data$organization[Billion_data$selfMade == 'FALSE' & is.na(Billion_data$organization)] <- "Inherited"
#Replace remaining true mission value with Unknown
Billion_data$organization[is.na(Billion_data$organization)] <- "Unknown"



```
-Title
```{r}
## Unknown values for their titles, might be because the person doesn't have a position at a company
# Fill missing values in 'title' with 'Unknown'
Billion_data$title[is.na(Billion_data$title)] <- "Unknown"
```

-residence State Region
```{r}
# Print the counts of unique values in 'residenceStateRegion'
table(Billion_data$residenceStateRegion)

# Print the counts of unique values in 'state'
table(Billion_data$state)
### Missing values in both attribute would be as result of the billionaires living outside of United State

```

```{r}
# Replace missing values in 'residenceStateRegion' with 'Outside US'
Billion_data$residenceStateRegion[is.na(Billion_data$residenceStateRegion)] <- "Outside US"

# Replace missing values in 'state' with 'Outside US'
Billion_data$state[is.na(Billion_data$state)] <- "Outside US"

# Print the updated counts
print(table(Billion_data$residenceStateRegion))
print(table(Billion_data$state))
```

-City
```{r}
## Probably could have used a method to fill these with other cities that reside in the same country
Billion_data$city <- ifelse(is.na(Billion_data$city), Billion_data$country, Billion_data$city)
```


-Country
```{r}
## Using their country of citizenship as a replacement value for country is the best option here
Billion_data$country <- ifelse(is.na(Billion_data$country), Billion_data$countryOfCitizenship, Billion_data$country)


```

# Numberical Observation 
```{r}
## Unfortunately, it looks like this is just missing data. We'll use the mean of each, as they are national averages and most likely won't change that much
## Considering we're dealing with mostly developed countries.
# Assuming 'Billion_data' is your data frame

# Replace missing values in 'cpi_change_country' with the mean
Billion_data$cpi_change_country[is.na(Billion_data$cpi_change_country)] <- mean(Billion_data$cpi_change_country, na.rm = TRUE)

# Replace missing values in 'cpi_country' with the mean
Billion_data$cpi_country[is.na(Billion_data$cpi_country)] <- mean(Billion_data$cpi_country, na.rm = TRUE)

# Replace missing values in 'tax_revenue_country_country' with the mean
Billion_data$tax_revenue_country_country[is.na(Billion_data$tax_revenue_country_country)] <- mean(Billion_data$tax_revenue_country_country, na.rm = TRUE)

# Replace missing values in 'total_tax_rate_country' with the mean
Billion_data$total_tax_rate_country[is.na(Billion_data$total_tax_rate_country)] <- mean(Billion_data$total_tax_rate_country, na.rm = TRUE)

# Replace missing values in 'life_expectancy_country' with the mean
Billion_data$life_expectancy_country[is.na(Billion_data$life_expectancy_country)] <- mean(Billion_data$life_expectancy_country, na.rm = TRUE)

# Print the updated data frame
print(Billion_data)

```


-gdp_country
```{r}
# Remove '$' and ',' from 'gdp_country' and convert to numeric
Billion_data$gdp_country <- as.numeric(gsub("[\\$,]", "", Billion_data$gdp_country))

```

```{r}

# Print summary statistics for 'population_country' and 'gdp_country'

print(summary(Billion_data[c('population_country', 'gdp_country')]))
sd(Billion_data$population_country,na.rm = TRUE)
sd(Billion_data$gdp_country,na.rm = TRUE)

```
```{r}
# Replace missing values in 'population_country' with the median, since the standard deviation is greater than the mean value
median_population <- median(Billion_data$population_country, na.rm = TRUE)
Billion_data$population_country[is.na(Billion_data$population_country)] <- median_population

# Replace missing values in 'gdp_country' with the median
median_gdp <- median(Billion_data$gdp_country, na.rm = TRUE)
Billion_data$gdp_country[is.na(Billion_data$gdp_country)] <- median_gdp

```
```{r}
# Print summary statistics for the specified columns
print(summary(Billion_data[c('gross_primary_education_enrollment_country', 'gross_tertiary_education_enrollment')]))
sd(Billion_data$gross_primary_education_enrollment_country,na.rm = TRUE)
sd(Billion_data$gross_tertiary_education_enrollment,na.rm = TRUE)


```

```{r}
# Replace missing values in 'gross_primary_education_enrollment_country' with the mean
mean_primary_education <- mean(Billion_data$gross_primary_education_enrollment_country, na.rm = TRUE)
Billion_data$gross_primary_education_enrollment_country[is.na(Billion_data$gross_primary_education_enrollment_country)] <- mean_primary_education

# Replace missing values in 'gross_tertiary_education_enrollment' with the mean
mean_tertiary_education <- mean(Billion_data$gross_tertiary_education_enrollment, na.rm = TRUE)
Billion_data$gross_tertiary_education_enrollment[is.na(Billion_data$gross_tertiary_education_enrollment)] <- mean_tertiary_education


```

```{r}
# Replace missing values in 'longitude_country' with the mean
mean_longitude <- mean(Billion_data$longitude_country, na.rm = TRUE)
Billion_data$longitude_country[is.na(Billion_data$longitude_country)] <- mean_longitude

# Replace missing values in 'latitude_country' with the mean
mean_latitude <- mean(Billion_data$latitude_country, na.rm = TRUE)
Billion_data$latitude_country[is.na(Billion_data$latitude_country)] <- mean_latitude

```

Age
```{r}
print(summary(Billion_data['age']))
# Replace missing values in 'age' with the mean
mean_age <- mean(Billion_data$age, na.rm = TRUE)
Billion_data$age[is.na(Billion_data$age)] <- mean_age

```

```{r}
# Create a new column 'first_name' by splitting the 'name' column
Billion_data$firstName <- sapply(strsplit(as.character(Billion_data$personName), " "), function(x) x[1])


```


```{r}

# Drop 'birthmonth' and 'birthdate' columns
Billion_data <- Billion_data[, !names(Billion_data) %in% c('birthDate','birthYear','birthMonth','birthDay')]

```

```{r}
# Assuming 'billionaires' is your data frame

# Load required library
#install.packages("plotly")

library(plotly)

# Create the scatter plot on a world map
fig1 <- plot_ly(
  type = 'scattergeo',
  mode = 'markers',
  lon = ~Billion_data$longitude_country,
  lat = ~Billion_data$latitude_country,
  marker = list(
    size = 8,
    opacity = 0.7,
    color = ~Billion_data$finalWorth,
    colorscale = 'Rainbow',
    colorbar = list(title = 'Final Worth')
  )
)

# Update the map projection and layout
fig1 <- fig1 %>% 
  layout(
    geo = list(projection = list(type = 'natural earth')),
    title = 'Billionaire Wealth Distribution on World Map'
  )

# Show the plot
fig1

```


```{r}
ggplot(Billion_data, aes(x=finalWorth)) + 
  geom_histogram(binwidth = 5,color="black")

```

```{r}
ggplot(Billion_data, aes(x=age)) + 
  geom_histogram(binwidth = 5,color="black")
```
```{r}
# Load the ggplot2 library
library(ggplot2)

# Create a bar plot for 'gender'
ggplot(Billion_data, aes(x = gender, fill = gender)) +
  geom_bar(color = "black", alpha = 0.7) +
  labs(title = "Distribution of Gender", x = "Gender", y = "Frequency") +
  theme_minimal()

# Comparing male and female final worth
Billion_data |> 
  ggplot(aes(finalWorth, fill = gender)) + 
  geom_density(alpha = 0.8) +
  scale_x_log10()

# Comparing selfmade billionaires
Billion_data |> 
  ggplot(aes(finalWorth, fill = selfMade)) + 
  geom_density(alpha = 0.8) +
  scale_x_log10()

Billion_data |> 
  ggplot(aes(x = industries, finalWorth)) +
  geom_histogram(aes(fill = selfMade), stat = "identity") +
  coord_flip() +
  facet_wrap(~ gender)
```
```{r}
library(dplyr)


df_grp_reg_cat = Billion_data %>% group_by(Billion_data$country, category) %>%
				summarise( mean_finalworth = mean(finalWorth), 
							.groups = 'drop')

View(df_grp_reg_cat)

```


```{r}
# Load the required libraries
library(plotly)
library(dplyr)

# Check for and remove duplicate levels in 'category'
df_grp_reg_cat <- df_grp_reg_cat %>%
  filter(!duplicated(category))

# Reorder 'category' based on the sum of 'mean_finalworth'
df_grp_reg_cat$category <- factor(df_grp_reg_cat$category, levels = df_grp_reg_cat$category[order(-df_grp_reg_cat$mean_finalworth)])

# Create an interactive bar plot for total final worth by category
fig <- plot_ly(df_grp_reg_cat, x = ~category, y = ~mean_finalworth, type = 'bar', color = ~category) %>%
  layout(title = "Total Final Worth by Category",
         xaxis = list(title = "Category"),
         yaxis = list(title = "Total Final Worth"),
         showlegend = FALSE) %>%
  colorbar(title = "Category")

# Show the plot
fig
```

```{r}
subset_data1 <- df_grp_reg_cat[df_grp_reg_cat$`Billion_data$country` == "United States", ]

```
```{r}
# Reorder 'category' based on the sum of 'mean_finalworth'
subset_data1$category <- factor(subset_data1$category, levels = subset_data1$category[order(-subset_data1$mean_finalworth)])

# Create an interactive bar plot for total final worth by category
fig <- plot_ly(subset_data1, x = ~category, y = ~mean_finalworth, type = 'bar', color = ~category) %>%
  layout(title = "Avearge Final Worth by Category for united state",
         xaxis = list(title = "Category"),
         yaxis = list(title = "Total Final Worth"),
         showlegend = FALSE) %>%
  colorbar(title = "Category")

# Show the plot
fig
```


```{r}

impute_missing <- function(data) {
  for (col in colnames(data)) {
    if (is.numeric(data[, col])) {
      sd_col <- sd(data[, col], na.rm = TRUE)
      mean_col <- mean(data[, col], na.rm = TRUE)

      if (sd_col < mean_col / 2) {
        data[, col][is.na(data[, col])] <- mean_col
      } else {
        data[, col][is.na(data[, col])] <- median(data[, col], na.rm = TRUE)
      }
    }
  }

  return(data)
}

```

```{r}
impute_missing <- function(data) {
  impacted_cols <- NULL

  for (col in colnames(data)) {
    if (is.numeric(data[, col])) {
      # Convert empty strings to NA before imputation
      data[, col][data[, col] == ""] <- NA

      sd_col <- sd(data[, col], na.rm = TRUE)
      mean_col <- mean(data[, col], na.rm = TRUE)

      if (sd_col < mean_col) {
        data[, col][is.na(data[, col])] <- mean_col
      } else {
        data[, col][is.na(data[, col])] <- median(data[, col], na.rm = TRUE)
      }

      impacted_cols <- c(impacted_cols, col)
    }
  }

  return(list(data = data, impacted_cols = impacted_cols))
}



```


```{r}
impute_missing(Billion_data)
colSums(is.na(Billion_data))

```

```{r}
# Count missing values in each column
missing_values <- colSums(is.na(Billion_data))
# Create a data frame with column names and their corresponding missing counts
missing_df <- data.frame(column = names(missing_values), missing_count = missing_values)
# Sort the data frame by missing counts in descending order
missing_df <- missing_df[order(-missing_df$missing_count), ]
```

